---
title: "Customer Churn - EDA"
author: "Christopher Hughes"
date: "30/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(caret)
library(readr)
# Set seed for reproducibility
set.seed(123)
```

## 1. Define Problem

Using the dataset to assess the important features that management should consider in developing customer retention programme.

We focussed on two main questions:
* Can we accurately predict which customers will churn?
* Which factors heavily relate churn and can be used in a save strategy?

With sub questions of:
* How the model performs
* The financial savings from its implementation

## 2. Exploratory Data Analysis

```{r}
aia <- readr::read_csv('../data/AIA_Churn_Modelling_Case_Study.csv')
knitr::kable(head(aia)[,1:8])
```

### Missing values

Review Missing Values by Feature. We see that TotalCharges is missing for 11 rows of the data. Customers with a `Tenure` of 0 have a missing `TotalCharge`. These have been imputed with 0.

```{r}
missing = colSums(is.na(aia))
data.frame(Feature = names(missing), Missing = missing) %>%
  arrange(desc(Missing)) %>%
  head %>%
  knitr::kable()
```

### Tenure Costs and TotalCharges

`Tenure` and `TotalCharge` are highly correlated. `Tenure` x `MonthlyCharges` is not equal to `TotalCharges`. An explanation is outside of this dataset. It could be: changes in plans, fines, changes in price of plans.

```{r}
aia %>%
  mutate(tenMon = tenure*MonthlyCharges,
         diff = (tenMon-TotalCharges)) %>%
  select(tenure, MonthlyCharges, TotalCharges, tenMon, diff, Churn) %>%
  head %>%
  knitr::kable()
```

## 3.Data Visualisation

The data is a mix of categorical and numeric variables. Data types converted for modelling and plotting.

```{r, message=FALSE}
aia <- aia %>%
  select(-customerID) %>%
  mutate(gender = factor(gender),SeniorCitizen = factor(SeniorCitizen),
         Partner = factor(Partner), Dependents = factor(Dependents),gender = factor(gender),
         PhoneService = factor(PhoneService),MultipleLines = factor(MultipleLines),
         InternetService = factor(InternetService),OnlineSecurity = factor(OnlineSecurity),
         DeviceProtection = factor(DeviceProtection),TechSupport = factor(TechSupport),
         StreamingTV = factor(StreamingTV),StreamingMovies = factor(StreamingMovies),
         Contract = factor(Contract),PaperlessBilling = factor(PaperlessBilling),
         PaymentMethod = factor(PaymentMethod), Churn = factor(Churn)) %>%
  mutate(TotalCharges = if_else(is.na(TotalCharges), 0, TotalCharges))
```

### Categorical Variables

Gender is an uninteresting variable and will be dropped from modelling

```{r, warning=FALSE}
aia %>%
  select(-c(tenure, MonthlyCharges, TotalCharges)) %>%
  gather(key="key", value="value", -Churn) %>%
  drop_na() %>%
  ggplot(aes(x = value, fill = Churn)) +
  geom_bar(alpha = 0.5) +
  facet_wrap(~key, scales = "free")
```

### Continuous Variables

```{r}
aia %>%
  select(c(tenure, MonthlyCharges, TotalCharges, Churn)) %>%
  gather(key="key", value="value", -Churn) %>%
  drop_na() %>%
  ggplot(aes(x = value, fill = Churn)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~key, scales = "free")
```

High correlation between `tenure` and `TotalCharges`. 

```{r}
aia %>%
  select(c(tenure, MonthlyCharges, TotalCharges)) %>%
  drop_na() %>%
  cor() %>%
  corrplot::corrplot(order = "hclust")
```

High paying customers tend to churn over time

```{r}
aia %>%
  drop_na() %>%
  ggplot() +
  geom_point(aes(x=tenure, y = TotalCharges, colour = Churn)) +
  ggtitle('High paying customers tend to churn over time')
```

#### Spread of data across tenure

There do not appear to be outliers in the `MonthlyCharges` when viewed by `tenure`.

```{r}
aia %>%
  select(c(tenure, MonthlyCharges)) %>%
  drop_na() %>%
  ggplot() +
  geom_boxplot(aes(group=tenure, y=MonthlyCharges))
```

